Web Authn Reference Implementation

### 2019-02-04 doc/paper/paper.tex
final

By using certified authenticators, \gls{fido2} makes sure that a user needs to have access to his authenticator and be present when the authentication is triggered. With this, an attacker could still steal the physical authenticator and authenticate with it. However, this kind of physical attack can't be used to steal the authenticator for a large group of people. Additionally, once the user notices that his authenticator has been stolen, he can inform the service provider and disable this authenticator for future authentications. Moreover, to secure authenticators that are left in the USB slot of a machine, \gls{fido2} requires the user to be present during the authentication. With this, an attacker can't merely put some malware on the user's computer and trigger the login.

@@

\begin{figure}[h]
  \includegraphics[width=12cm]{img/u2fcompat-getassertion.png}
  \centering
  \caption{Authenticator Data}
  \label{fig:authenticationMapping}
\end{figure}

@@

The create registration step consists of creating a \gls{json} which represents a valid challenge. This JSON is sent to the browser where it is handed to the call `navigator.create()'. The response is sent to the /response endpoint in the replying party.

@@

\lstinputlisting[language=json, caption=Registration JSON, label=lst:createRegistration]{json/createRegistration.json}

\begin{description}
  \item[challenge] \hfill \\ The challenge to be signed by the Authenticator. Must be unique for each call and base 64 encoded.
  \item[fidoResponse] Should be direct. 
  \item[rp.id] \hfill \\ Must reflect the domain over which you communicate
  \item[user.id] Must be unique and random. 
  \item[pubKeyCredParams.type] Should be public-key. 
  \item[pubKeyCredParams.alg] Must be a valid \gls{cose} algorithm. \cite{cose}
  \item[attestation] Should be direct.
  \item[timeout] Should be defined.   
\end{description}

@@

\lstinputlisting[language=json, caption=Create Response JSON, label=lst:create_response]{json/createResponse.json}

@@

\lstinputlisting[language=json, caption=Client Data JSON Parsed, label=lst:clientDataParsed]{json/clientDataParsed.json}

@@

\lstinputlisting[language=json, caption=Create Authentication JSON, label=lst:createAuthentication]{json/createAuthentication.json}

@@

\lstinputlisting[language=json, caption=Attestation Response JSON, label=lst:attestationResponse]{json/attestationResponse.json}

@@

\lstinputlisting[language=json, caption=Attestation Client Data JSON, label=lst:attestationClientData]{json/attestationClientData.json}

@@

\begin{figure}[ht]
  \includegraphics[width=12cm]{img/authenticatorResponseData.png}
  \centering
  \caption{Authenticator Data}
  \label{fig:authenticatorData}
\end{figure}

\begin{figure}[ht]
  \includegraphics[width=12cm]{img/authenticatorData.png}
  \centering
  \caption{Attested Credential Data}
  \label{fig:credentialData}
\end{figure}

>>>>

In \gls{fido2} the Authenticator generally receives a challenge when it is accessed. This challenge comes from the replying party which wants to authenticate a user. With the challenge, the replying party sends the domain name where it is available. The authenticator can then check if that is truly the domain where the user is accessing the site. Should this be valid, the authenticator creates a response by signing the challenge and the host. The replying party can then during validation check whether the challenge and the domain contain the requested values and whether the signature is correct. With this, a \gls{mitm} attacker can't forward a challenge from a replying party to an authenticator by using a spoofed page. The user and the replying party can both be assured that they are talking to the correct party. To assure this \gls{tls} is mandatory for each session. \cite{yubico:whatIsFido2}

By using certified authenticators, \gls{fido2} makes sure that a user needs to have access to his authenticator and be present when the authentication is triggered. With this, an attacker could still steal the physical authenticator and authenticate with it. However, this kind of physical attack can't be used to steal the authenticators for a large group of people. Additionally, once the user notices that his authenticator has been stolen, he can inform the service provider and disable this authenticator for future authentications. Moreover, by requiring the user to be present during the authentication authenticators that are left plugged into the machine are also secured. An attacker remotely triggers the authentication even if the authenticator is left in the machine for a long time.

@@

\begin{figure}[h]
  \includegraphics[width=16cm]{img/u2fcompat-getassertion.png}
  \centering
  \caption{Authenticator Data}
  \label{fig:authenticationMapping}
\end{figure}

@@

The create registration step consists of creating a \gls{json} which represents a valid challenge. This \gls{json} is sent to the browser where it is handed to the call `navigator.create()'. The response is sent to the /response endpoint in the replying party.

@@

\lstinputlisting[language=json, caption=Registration \gls{json}, label=lst:createRegistration]{json/createRegistration.json}

\begin{description}
  \item[challenge] \hfill \\ The challenge to be signed by the Authenticator. Must be unique for each call and \gls{base64} encoded.
  \item[fidoResponse] Should be direct. 
  \item[rp.id] \hfill \\ Must reflect the domain over which you communicate
  \item[user.id] Must be unique and random. 
  \item[pubKeyCredParams.type] Should be public-key. 
  \item[pubKeyCredParams.alg] Must be a valid \gls{cose} algorithm. \cite{cose}
  \item[attestation] Should be direct.
  \item[timeout] Should be defined.   
\end{description}

@@

\lstinputlisting[language=json, caption=Create Response \gls{json}, label=lst:create_response]{json/createResponse.json}

@@

\lstinputlisting[language=json, caption=Client Data \gls{json} Parsed, label=lst:clientDataParsed]{json/clientDataParsed.json}

@@

\lstinputlisting[language=json, caption=Create Authentication \gls{json}, label=lst:createAuthentication]{json/createAuthentication.json}

@@

\lstinputlisting[language=json, caption=Attestation Response \gls{json}, label=lst:attestationResponse]{json/attestationResponse.json}

@@

\lstinputlisting[language=json, caption=Attestation Client Data \gls{json}, label=lst:attestationClientData]{json/attestationClientData.json}

@@

\begin{figure}[ht]
  \includegraphics[width=16cm]{img/authenticatorResponseData.png}
  \centering
  \caption{Authenticator Data}
  \label{fig:authenticatorData}
\end{figure}

\begin{figure}[ht]
  \includegraphics[width=16cm]{img/authenticatorData.png}
  \centering
  \caption{Attested Credential Data}
  \label{fig:credentialData}
\end{figure}
